Statistical Theory\\ Chess Dataset Analysis

### 2024-09-30 Article/Article.tex
Updated src
Updated Article

@@

The most comprehensive way to analyze a dataset is through the classification of its distribution. To achieve this, we examine the moments of the distribution for the variable "Turns" as follows:
\[Mean: 61.96255029402662\]
\[Standard Deviation: 33.731730929355365\]
\[Skewness: 0.9052618256476783\]
\[Kurtosis: 1.4921324651335777\]

@@

Despite this improvement, the Generalized Gamma distribution does not sufficiently account for the excess kurtosis. Therefore, we combined it with a Poisson distribution to create a mixture model that captures both the skewness and kurtosis properties. The results of this hybrid model are illustrated in Figure~\ref{fig:gam_poi_fit}.

>>>>

\begin{abstract}
This study explores the relationship between \textbf{player ratings} and \textbf{chess game duration} using a dataset of online games. We aim to determine whether \textbf{higher-rated players} tend to play longer or shorter games and investigate other factors influencing game length, such as \textbf{time controls} and \textbf{game outcomes}. A cleaned dataset was analyzed through statistical techniques, including the \textbf{Kolmogorov-Smirnov test} for normality, the \textbf{Mann-Whitney U test} for differences in game length between rating groups, and \textbf{correlation analysis}. Results indicate that \textbf{higher-rated players} generally engage in longer games, while greater disparities in player ratings lead to shorter matches. Additionally, \textbf{draws} tend to prolong game duration compared to decisive outcomes. The game's turn count distribution shows positive skewness and excess kurtosis, leading us to model it with a \textbf{Gamma-Poisson mixture}. This model effectively captures the underlying distribution characteristics, providing further insights into the statistical patterns governing chess match dynamics.
\end{abstract}

@@

The most comprehensive way to analyze a dataset is through the classification of its distribution. To achieve this, we examine the moments of the distribution for the variable "Turns" as follows:
\[Mean: 61.962\]
\[Standard Deviation: 33.731\]
\[Skewness: 0.9052\]
\[Kurtosis: 1.492\]

@@

Despite this improvement, the Generalized Gamma distribution does not sufficiently account for the excess kurtosis. Therefore, we combined it with a Poisson distribution to create a mixture model that captures both the skewness and kurtosis properties. The results of this hybrid model are illustrated in Figure~\ref{fig:gam_poi_fit}. This fit actually performs amazingly for the 2 sample KS test, even sometimes beating it, with a statistic of $0.01429$ and a p value of $0.073$, which is above the arbitrary threshold of 0.05.
